The tracing of trade: exploring the patterns of olive oil production and distribution from Roman \textit{Baetica}

### 2020-11-04 Paper/jaccardbaetica.tex
add review VML and git repository

%\title{The markings of the trade: exploring the patterns of olive oil production in Roman Baetica}
\title{The tracing of trade: exploring the patterns of olive oil production and distribution from Roman Baetica}

@@

The structure and processes associated with this massive olive oil production have been extensively discussed over the last decades\citep{rodriguez_economioleicola_1977, Chic_hispania_1997,millet_anforas_1998}. 
However, the lack of written records has made possible to detect any indication of patterns in the olive oil market.

@@

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/general_map}
\caption{Overview of the sites analysed in this paper. The color defines the three different regions under study (\textit{Baetica} as producer; \textit{Germania} and \textit{Britannia} as consumers) while the size of each dot highlights the number of Dressel-20 amphoric stamps found on each of these sites}

@@

In the case of \textit{Baetica} province, we want to identify the role of the stamps in the organisation of the workshop; in Roman provinces, our aim is detecting groups of stamps concentrated in an area or if some groups have an important role for the exportation of olive oil in those provinces. This economic connection could be identified by different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamps in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific province.

@@

The stamp codes are interpreted based on three main ideas: content (olive oil), context (amphora workshop) and subject (individuals involved in the production). On the one hand, it seems that stamps could have been identified as the landowner of the olive groves \citep{rodriguez_economioleicola_1977}. On the other hand, they could also belong to the owner of the workshops were the amphorae were made or even to a group of amphora workers~\citep{berni_millet_epigrafianforica_2008}. 
%In any case, the use of these stamps is a good proxy to explore the system of olive oil production and trade developed in Baetica.

@@

The consumption of olive oil would experience a progressive slowdown from the third century A.D. onwards. This date matches a change in market strategy in the Empire. This a gradual decrease on \textit{Baetican} olive oil exports can be observed as a decrease in the amount of Dressel 20 found in excavated sites as this typology is replaced by other types of amphorae from other places \citep{rodriguez1991aceite,millet_anforas_1998}.

@@

The presence of the Roman army encouraged the exchange in the province as shown by the arrival of this product both civil settlement and military sites with a mayor concentration at the German \textit{limes}.

@@

The type of distribution is also frequent in amphora production where we observed a major concentration of the number of stamps in few workshops\citep{bayesian_2018,coto-sarmiento_identifying_2018}.

@@

The distribution of code stamps for each \textit{conventus} can be seen in Figure~\ref{frequency}. The majority of stamps are concentrated in \textit{Hispalis} (574 code stamps) while \textit{Corduba} and \textit{Astigi} have roughly half this sample size (267 and 146 code stamps). The workshops of the three \textit{conventus} show a similar pattern on the stamp frequency distribution with the exceptions of 2 large workshops in \textit{Hispalis} (La Catria and Arva). On these two workshops a comparatively large amount of code stamps were found (29 different code stamps on each of them)

@@

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/germania}
\caption{Sites in Germania where Dressel-20 amphoric stamps have been found. Each dot is a site while its size shows the sample size of stamps found on the site. Most sites are located along the German \textit{limes}}
\label{germania}
\end{figure}

@@

Considering our dataset as a non-uniform sample, this method provides a useful tool to handle large samples with different sizes and diversity \citep{wolda_similarity_1981}. Morisita-Horn index gives a value between 0 (sites have exact stamp codes and frequency distribution) and 1 (complete difference between stamp codes). To apply Morisita-Horn we first calculated the number of times that each stamp code appeared in an amphora workshop. 
This method allowed us to bear in mind the number of frequency of stamps for each workshop.

@@

Results for the consumption regions display similar patterns than the ones found for the production area of \textit{Baetica}. The most important difference is that the similarity values are lower as can be seen in Figure~\ref{britmap}

@@

This could be explained such as batch systematic organisation between potters \citep{juanmorostesis}. This work method could have allowed to potters to stamp amphorae in order to identify the production for batches for posterior commercialization. High production of amphorae had to maintain a specialize system organization to control this production \citep[104]{juanmorostesis}.

%This method allowed to potters organises the production with the batch stamping for the posterior commercialisation
%\xavi{y esto para qué serviría? debería explicarse un poco más}.

@@

This evidence means that it is unknown whether some production centres went to one province or another or, at least, that they can be clearly reflected in the data with a greater similarity in the amphoric stamps. Thus, production centres could have distributed randomly olive oil both \textit{Britannia} and \textit{Germania}. Neither we do not detect production centres dedicated to the distribution of olive oil in a certain province.

@@

This research was partly funded by the European Research Council Advanced Grant EPNet (340828). We are grateful to Simon Carrignon, Juan Moros and Ignacio Morer for their useful suggestions.  
All data has been analysed and conducted in R program version 3.2.4, using the packages \textit{vegan} \citep{oksanen_vegan_2007}, \textit{ggplot2} \citep{ggplot2:_2016}. Source and code are available at \maria{incluir github}.

>>>>

%\title{The markings of the trade: exploring the patterns of olive oil production in Roman Baetica}
\title{The tracing of trade: exploring the patterns of olive oil production and distribution from Roman \textit{Baetica}}

@@

The structure and processes associated with this massive olive oil production have been extensively discussed over the last decades\citep{rodriguez_economioleicola_1977, Chic_hispania_1997,millet_anforas_1998}. 
However, the lack of written records has not made possible to detect any indication of patterns in the olive oil market.

@@

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/general_map}
\caption{Overview of the sites analysed in this paper. The colour defines the three different regions under study (\textit{Baetica} as producer; \textit{Germania} and \textit{Britannia} as consumers) while the size of each dot highlights the number of Dressel 20 amphoric stamps found on each of these sites}

@@

In the case of \textit{Baetica} province, we want to identify the role of the stamps in the organisation of the workshop; in Roman provinces, our aim is detecting groups of stamps concentrated in an area or if some groups have an important role for the exportation of olive oil in those provinces. This economic connection could be identified by different aspects: a) correlation between spatial distance and the sites based on the idea that closer workshops concentrate similar amphoric stamps in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific province.

@@

The stamp codes are interpreted based on three main ideas: content (olive oil), context (amphora workshop) and subject (individuals involved in the production). On the one hand, it seems that stamps could have been identified as the landowner of the olive groves \citep{rodriguez_economioleicola_1977}. On the other hand, they could also belong to the owner of the workshops where the amphorae were made or even to a group of amphora workers~\citep{berni_millet_epigrafianforica_2008}. 
%In any case, the use of these stamps is a good proxy to explore the system of olive oil production and trade developed in Baetica.

@@

The consumption of olive oil would experience a progressive slowdown from the third century A.D. onwards. This date matches a change in market strategy in the Empire. This gradual decrease on \textit{Baetican} olive oil exports can be observed as a decrease in the amount of Dressel 20 found in excavated sites as this typology is replaced by other types of amphorae from other places \citep{rodriguez1991aceite,millet_anforas_1998}.

@@

The presence of the Roman army encouraged the exchange in the province as shown by the arrival of this product both civil settlement and military sites with a major concentration at the German \textit{limes}.

@@

The type of distribution is also frequent in amphora production where we observed a major concentration of the number of stamps in few workshops~\citep{bayesian_2018,coto-sarmiento_identifying_2018}.

@@

The distribution of code stamps for each \textit{conventus} can be seen in Figure~\ref{frequency}. The majority of stamps are concentrated in \textit{Hispalis} (574 code stamps) while \textit{Corduba} and \textit{Astigi} have roughly half this sample size (267 and 146 code stamps). The workshops of the three \textit{conventus} show a similar pattern on the stamp frequency distribution with the exceptions of 2 large workshops in \textit{Hispalis} (La Catria and Arva). On these two workshops a comparatively large amount of code stamps were found (29 different code stamps on each of them).

@@

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/germania}
\caption{Sites in Germania where Dressel 20 amphoric stamps have been found. Each dot is a site while its size shows the sample size of stamps found on the site. Most sites are located along the German \textit{limes}}
\label{germania}
\end{figure}

@@

Considering our dataset as a non-uniform sample, this method provides a useful tool to handle large samples with different sizes and diversity \citep{wolda_similarity_1981}. Morisita-Horn index gives a value between 0 (sites have exact stamp codes and frequency distribution) and 1 (complete difference between stamp codes). To apply Morisita-Horn, we first calculated the number of times that each stamp code appeared in an amphora workshop. 
This method allowed us to bear in mind the number of frequency of stamps for each workshop.

@@

Results for the consumption regions display similar patterns than the ones found for the production area of \textit{Baetica}. The most important difference is that the similarity values are lower as can be seen in Figure~\ref{britmap}.

@@

This could be explained such as batch systematic organisation between potters \citep{juanmorostesis}. This work method could have allowed potters to identify batches by stamping amphorae for posterior commercialization.

%This work method could have allowed to potters to stamp amphorae in order to identify the production for batches for posterior commercialization.

High production of amphorae had to maintain a specialize system organization to control this production \citep[104]{juanmorostesis}.

%This method allowed to potters organises the production with the batch stamping for the posterior commercialisation

@@

This evidence means that it is unknown whether some production centres went to one province or another or, at least, that they can be clearly reflected in the data with a greater similarity in the amphoric stamps. Thus, production centres could have distributed randomly olive oil to both \textit{Britannia} and \textit{Germania}. Neither we do not detect production centres dedicated to the distribution of olive oil in a certain province.

@@

This research was partly funded by the European Research Council Advanced Grant EPNet (340828). We are grateful to Simon Carrignon, Juan Moros, Ignacio Morer and Víctor M. Lozano for their useful suggestions.  
Data have been analysed and conducted in R program version 3.2.4, using the packages \textit{vegan} \citep{oksanen_vegan_2007}, \textit{ggplot2} \citep{ggplot2:_2016}. Datasets and source code are freely available at https://github.com/Mcotsar/Baetica\_stamps.
